Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-18 forecast/paper/paper.tex
Fixed a reference, proofreading up to results.

% Please add a significance statement to explain the relevance of your work
\significancestatement{Forecasting is a critical element of the public health
  response to any fast-moving epidemic or pandemic.  Predicting the future
  spread some disease process, based on disease activity in the recent
  past---say, using what is known as autoregressive modeling in the time series
  literature---is a basic and widely-used method.  In this paper we study
  whether including auxiliary indicators as additional features in such an
  autoregressive model can be useful in COVID-19 case forecasting and hotspot
  prediction.  We find that indicators based on de-identified medical insurance 
  claims, self-reported symptoms through online surveys, and COVID-related
  Internet search activity all provide a nontrivial boost in accuracy, though
  their utility generally depends the pandemic's local dynamics (upswing,
  downswing, or flat period) at prediction time.
}

@@

\begin{abstract}
Please provide an abstract of no more than 250 words in a single paragraph.
Abstracts should explain to the general reader the major contributions of the
article. References in the abstract must be cited in full within the abstract
itself and cited in the text. 
\end{abstract}

@@

We consider prediction of future COVID-19 case rates or case hotspots (to be
defined precisely shortly).  By case rate, we mean the case count per 100,000
people (the standard in epidemiology).  We use reported case data aggregated by 
JHU CSSE \cite{Dong:2020}, which is accessible through the COVIDcast API
\cite{CovidcastAPI}, like the auxiliary indicators that we use to supplement the
basic purely autoregressive models.

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey CLI in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community that is sick, based on Delphi's surveys of
  Facebook users, called the COVID Trends and Impact Survey. 
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
  COVID-19 related Google search volume for queries relating to anosmia or
  ageusia (loss of smell or taste), based on Google's Search Trends data set. 
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

For more details on the five indicators (including how these are precisely
computed from the underlying data streams) we refer to
\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
which documents all of the signals in the COVIDcast API, and our companion paper
on the API and database. For CTIS in particular, we refer to our companion paper
on this survey. For the Google Search Trends data set, see
\cite{GoogleSymptoms}, and \cite{Klopfen:2020, Vaira:2020} for articles on the
relevance of anosmia or ageusia (loss of smell or taste) to COVID.

As for geographic resolution, we consider the prediction of COVID-19 case rates
and hotspots aggregated at the level of an individual \textit{hospital referral
  region} (HRR). HRRs correspond to groups of counties in the United States
within the same hospital referral system. The Dartmouth Atlas of Healthcare 
Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
characteristics. They are contiguous regions such that most of the hospital
services for the underlying population are performed by hospitals within the
region. Each HRR also contains at least one city where major procedures
(cardiovascular or neurological) are performed. The smallest HRR has a
population of about 120,000. While some are quite large (such as the one
containing Los Angeles, which has about 10 million people), generally HRRs 
are much more homogenous in size than the (approximately) 3200 counties,  
and they serve as a nice middle ground in between counties and states.

HRRs, by their definition, would be most relevant for forecasting hospital
demand.  We have chosen to focus on cases (forecasting and predicting
hotspots) at the HRR level since the indicators considered should be more 
useful in predicting case activity versus hospital demand, since the latter is
intuitively more closely connected (in time) to the events that are measured by
the given five indicators. Predicting case rates (and hotspots) at the HRR level
is still a reasonable goal in its own right; and moreover, it could be used to
feed predicted case information into downstream hospitalization models.

@@

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 1, 2020, we train models 
using only the data that would have been available to us \textit{as of}
September 1 (imagine that we can ``rewind'' the clock to September 1, and query
the COVIDcast API to get the latest data it would have had available at that
point in time.)  This is made possible by the COVIDcast API's comprehensive data   
versioning system (described in more detail in our companion paper).  We also
use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
process of training arbitrary prediction models over a sequence of prediction
dates, by constructing the proper sequence of vintage training data sets.

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
  \caption{Revision behavior for two signals in the HRR containing New York City
    HRR.  Each colored line corresponds to the data as reported on a particular
    date (\textit{as of} dates varying from September 28 through October
    19). The left panel shows the DV-CLI signal, which was regularly revised
    throughout the period, though the effects fade as we look further back in   
    time. In contrast, the right panel shows case rates reported by JHU CSSE,
    which remain ``as reported'' on September 28, with a huge spike towards the 
    end of this period, until a major correction is made on October 19, which
    brings down this spike and affects all prior data as well.}
  \label{fig:vintage}
\end{figure}

Vintage training data means different things, in practice, for different
signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
DV-CLI, are most often 3-5 days latent, and subject to a considerable but
regular degree of revision or ``backfill'' after their initial publication date.
The survey-based signal, CTIS-CLI-in-community, is 2 days latent, and rarely
undergoes any revision at all.  The target variable itself, reported COVID-19
case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
(sometimes clear anomalies) after initial publication.  Compared to the revision
pattern in the medical claims signals, which are much more systematic in nature,
revisions in case reports can be highly erratic. Big spikes or other anomalies
can occur in the data as reporting backlogs are cleared, changes in case
definitions are made, etc. Groups like JHU CSSE then work tirelessly to correct
such anomalies after first publication (e.g., they will attempt to
back-distribute a spike when a reporting backlog is cleared, by working with a
local authority to figure out how this should best be done), which can result
in very nontrivial and often unpredictable revisions.  See
Figure~\ref{fig:vintage} for an example.

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then.  Furthermore, the
latency of the signal was always at least one week through 2020.  However, this
signal is never revised after initial publication (confirmed via personal
communication with the Google team that produces this signal) and furthermore
the latency of the signal is not an unavoidable property of the data type, so we
simply use finalized signal values it, with zero latency, in our analysis.

@@

To fix notation, let $Y_{\ell,t}$ denote the 7-day trailing average of COVID-19
case incidence rates in location (HRR) $\ell$ and at time (day) $t$.  To be
clear, this is the number of new daily reported cases per 100,000
people, averaged over the 7-day period $t-6, \ldots t$. The first task we
consider---\textit{forecasting}---is to predict $Y_{\ell,t+a}$ for each
``ahead'' value $a=7,\ldots,21$.  The second task---\textit{hotspot
  prediction}---is to predict a binary variable defined in terms of the relative
change of $Y_{\ell,t+a}$ (relative to its value one week prior,
$Y_{\ell,t+a-7}$), again for each $a=7,\ldots,21$.

Why do we define the response variables via 7-day averaging? The short answer 
is robustness: averaging stabilizes the case time series, and accounts for 
uninteresting artifacts like weekend-weekday differences in the series.  Note
that we can of course equivalently view this (equivalent up to a constant
factor) as predicting the HRR-level COVID-19 case incidence rate \textit{summed}
over some 7-day period in the future, and predicting a binary variable derived
from this.

@@

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data, and lagged values
of signals for features as we will detail shortly) June 9, 2020 for forecasting,
and June 16, 2020 for hotspot prediction. (The bottleneck here is the
CTIS-CLI-in-community signal, which does not exist before early April 2020, when
the survey was first launched). The end date was chosen again with a
consideration to align both tasks as best as possible, and because few hotspots
exist post December 31, 2020, due to the general and gradual decline of the
pandemic in 2021.

@@

Recall $Y_{\ell,t}$ denotes the 7-day trailing average of COVID-19 case
incidence rates in location $\ell$ and at time $t$.  Separately for each
$a=7,\ldots,21$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
simple probabilistic forecasting model of the form:     
\begin{equation}
\label{eq:forecast_ar}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, s \leq t)  
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
\end{equation}
This model uses current case rates, and the case rates 7 and 14 days ago, in
order to predict (the quantiles of) case rates in the future.  We consider a
total of 7 quantile levels (chosen in accordance with the county-level quantile
levels suggested by the COVID-19 Forecast Hub),      
\begin{equation}
\label{eq:quantile_levels}
\tau \in \{0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975 \}.
\end{equation}
We fit \eqref{eq:forecast_ar} using \textit{quantile regression}
\cite{Koenker:1978, Koenker:2005, Koenker:2006} separately for each $\tau$, 
using data from all 306 HRRs, and within each HRR, using the most recent 21 days
of training data.  This gives us 6,426 training samples for each quantile 
regression problem.

In addition to this pure autoregressive model, we also consider five
probabilistic forecasting models of the form:  
\begin{multline}
\label{eq:forecast_ar_x}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t)  
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
\end{multline}
where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
CHNG-COVID, CTIS-CLI-in-community, DV-CLI, or Google-AA---at location $\ell$ and
time $t$. Note that we apply the same lags (current value, along with the values
7 and 14 days ago) for the auxiliary indicators as we do for the case
rates. Training then proceeds just as before: we use the same 7 quantile levels
in \eqref{eq:quantile_levels}, and fit quantile regression separately for each 
level $\tau$, using data from all 306 HRRs and a trailing window of 21 days of
training data.

At prediction time, in order to deal with possible crossing violations (a
predicted quantile at level $\tau$ exceeds a predicted quantile at level $\tau'
> \tau$), we apply a simple post-hoc sorting.  See Figure~\ref{fig:trajectory}
for an example forecast.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
\caption{Forecast from an autoregressive model for the New York City HRR made 
  on October 15 (vertical line).  The fan displays 50\%, 80\% and 95\% intervals
  while the orange curve shows the median forecast. The black curve represents
  ``finalized'' data, as reported in May 2021.} 
\label{fig:trajectory}
\end{figure}

@@

We treat hotspot prediction as a binary classification problem and use a setup 
altogether quite similar to the forecasting setup described previously.
Separately for each $a=7,\ldots,21$, to predict $Z_{\ell,t+a}$, we consider a
simple logistic model:
\begin{equation}
\label{eq:hotspot_ar}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, s \leq t) \big) 
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
\end{equation}
where $\logit(p) = \log(p/(1-p))$, the log-odds of $p$.

In addition to this pure autoregressive model, we also consider five
logistic models of the form:  
\begin{multline}
\label{eq:hotspot_ar_x}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t) \big)
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit the above models,
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x}, using logistic regression,   
pooling all 306 HRRs and using a 21-day trailing window for the training data.

@@

For forecasting, we evaluate the probabilistic forecasts produced by the
quantile models in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x} using 
\textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
\cite{Gneiting:2007}.  WIS is a proper score, meaning that its expectation
is minimized by the population quantiles of the target variable.  The use of WIS
in COVID-19 forecast scoring is discussed in \cite{Bracher:2021}; WIS is also
the main evaluation metric used in the COVID-19 Forecast Hub.

WIS is typically defined for quantile-based forecasts where the quantile levels
are symmetric around 0.5.  This is the case for our choice in
\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the case of
symmetric quantile levels, this is equivalent to a collection of central
prediction intervals $(\ell_\alpha, u_\alpha)$, parametrized by an exclusion
probability $\alpha$. The WIS of the forecaster $F$, evaluated at the target
variable $Y$, is defined by:
\begin{equation}
\label{eq:wis_intervals}
\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
\end{equation}
where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
smallest distance between $a$ and an element of $S$).  Note that, corresponding 
to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
\{0.05, 0.2, 0.5, 1\}$, resulting in 4 terms in the above sum.  By 
straightforward algebra, it is not hard to see WIS has an alternative 
representation in terms of the predicted quantiles themselves:
\begin{equation}
\label{eq:wis_quantiles}
\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
\end{equation}
where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
presentation, because the score nicely decouples into a ``sharpness'' component
(first term in each summand) and ``under/overprediction'' component (second term
in each summand).  The second form given in \eqref{eq:wis_quantiles} is
especially noteworthy in our current study as it reveals WIS is the same as the
quantile regression loss that we use to train our forecasting models (i.e., our
models are fit to optimize WIS averaged over the training set).

For hotspot prediction, we evaluate the probabilistic classifiers produced by
the logistic models in \eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using the 
area under the curve (AUC) of their true positive versus false positive rate
curve (which is traced out by varying the discrimination threshold).

The primary aggregation scheme that we will use in model evaluation and
comparisons will be to average WIS per forecaster per ahead value $a$, over all
forecast dates $t$ and locations $\ell$; and similarly, to compute AUC per
classifier and per ahead value $a$, over all forecast dates $t$ and locations
$\ell$.

>>>>

% Please add a significance statement to explain the relevance of your work
\significancestatement{Forecasting is a critical element of the public health
  response to any fast-moving epidemic or pandemic.  Predicting the future
  spread of some disease process, based on disease activity in the recent
  past---say, using what is known as autoregressive modeling in the time series
  literature---is a basic and widely-used method.  In this paper we study
  whether including auxiliary indicators as additional features in such an
  autoregressive model can be useful in COVID-19 case forecasting and hotspot
  prediction.  We find that indicators based on de-identified medical insurance 
  claims, self-reported symptoms through online surveys, and COVID-related
  internet% Capitalization is apparently
  % in question: https://en.wikipedia.org/wiki/Capitalization_of_Internet
  search activity all provide a nontrivial boost in accuracy, though
  their utility generally depends the pandemic's local dynamics (upswing,
  downswing, or flat period) at prediction time.
}

@@

\begin{abstract}
Reliable short-term, location-specific forecasts of a pandemic are a
key ingredient in effective public health decision making. Throughout the
COVID-19 pandemic, a plethora of real-time data signals were
collected, curated, and made publicly available by our team and data
partners, providing multiple views of the pandemic's progress. The
goal of this work is to measure the practical utility of these
auxiliary indicators from a forecasting perspective. We focus on five
indicators, derived from medical insurance claims data, web search
queries, and survey responses. For each indicator, we ask whether its
inclusion in a simple model leads to improved forecasting
performance relative to a similar model excluding it. Performance is considered across two tasks
at a forecasting horizon of up to about one month:
probabilistic forecasting of seven-day average case rates and binary
prediction of ``hotspots.'' Since the values of indicators are
commonly revised over time, we take special care to ensure that the data provided
to a forecaster is the version that would have been available at the
time that the forecast was made. We demonstrate how less careful
backtesting can lead to misleading retrospective evaluations. Our analysis
suggests consistent, although 
modest, gains in forecasting performance by using these indicators,
and further investigation suggests that these gains are related to periods when the
auxiliary signals are acting as ``leading 
indicators'' of case rates. 
\attn{Evaluating forecasters thoroughly and fairly is an important way that
we can learn from this pandemic to do better in the next. 251 :(}

% Please provide an abstract of no more than 250 words in a single paragraph.
% Abstracts should explain to the general reader the major contributions of the
% article. References in the abstract must be cited in full within the abstract
% itself and cited in the text.
\end{abstract}

@@

We consider prediction of future COVID-19 case rates or case hotspots (to be
defined precisely shortly).  By case rate, we mean the case count per 100,000
people (the standard in epidemiology).  We use reported case data aggregated by 
JHU CSSE \cite{Dong:2020}, which, like the auxiliary indicators that we use to supplement the
basic purely autoregressive models, is accessible through the COVIDcast API
\cite{CovidcastAPI}.

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey COVID-like illness in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community that is sick, based on Delphi's surveys of
  Facebook users, called the COVID Trends and Impact Survey. 
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
  Google search volume for queries relating to anosmia or
  ageusia (loss of smell or taste), based on Google's Search Trends data set. 
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

For more details on the five indicators (including how these are precisely
computed from the underlying data streams) we refer to
\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
which documents all of the signals in the COVIDcast API, and our companion paper
on the API and database. For CTIS in particular, we refer to our companion paper
on this survey. For the Google Search Trends data set, see
\cite{GoogleSymptoms}, and see \cite{Klopfen:2020, Vaira:2020} for justification of the
relevance of anosmia or ageusia to COVID-19 infection.

As for geographic resolution, we consider the prediction of COVID-19 case rates
and hotspots aggregated at the level of an individual \textit{hospital referral
  region} (HRR). HRRs correspond to groups of counties in the United States
within the same hospital referral system. The Dartmouth Atlas of Healthcare 
Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
characteristics. They are contiguous regions such that most of the hospital
services for the underlying population are performed by hospitals within the
region. Each HRR also contains at least one city where major procedures
(cardiovascular or neurological) are performed. The smallest HRR has a
population of about 125,000. While some are quite large (such as the one
containing Los Angeles, which has more than 10 million people), generally HRRs 
are much more homogenous in size than the (approximately) 3200 counties,  
and they serve as a nice middle ground in between counties and states.

HRRs, by their definition, would be most relevant for forecasting hospital
demand.  We have chosen to focus on cases (forecasting and predicting
hotspots) at the HRR level because the indicators considered should be more 
useful in predicting case activity rather than hospital demand. The \attn{former?} is
intuitively more contemporaneous to the events that are measured by
the given five indicators. Predicting case rates (and hotspots) at the HRR level
is still a reasonable goal in its own right; and moreover, it could be used to
feed predicted case information into downstream hospitalization models.

@@

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 28, 2020, we train models 
using only the data that would have been available to us \textit{as of}
September 28 (imagine that we can ``rewind'' the clock to September 28 and query
the COVIDcast API to get the latest data it would have had available at that
point in time.)  This is made possible by the COVIDcast API's comprehensive data   
versioning system (described in more detail in our companion paper).  We also
use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
process of training arbitrary prediction models over a sequence of prediction
dates, by constructing the proper sequence of vintage training data sets.

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
  \caption{This figure displays revision behavior for two signals in the HRR
    containing New York 
    City.  Each colored line corresponds to the data as reported on a particular 
    date (\textit{as of} dates varying from September 28 through October
    19). The left panel shows the DV-CLI signal, which was regularly revised
    throughout the period, though the effects fade as we look further back in   
    time. In contrast, the right panel shows case rates reported by JHU CSSE,
    which remain ``as reported'' on September 28, with a huge spike towards the 
    end of this period, until a major correction is made on October 19, which
    brings down this spike and affects all prior data as well.}
  \label{fig:vintage}
\end{figure}

Vintage training data means different things, in practice, for different
signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
DV-CLI, are most often 3-5 days latent, and subject to a considerable but
regular degree of revision or ``backfill'' after their initial publication date.
The survey-based signal, CTIS-CLI-in-community, is 2 days latent, and rarely
undergoes any revision at all.  The target variable itself, reported COVID-19
case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
after initial publication.  Compared to the revision
pattern in the medical claims signals, which are much more systematic in nature,
revisions in case reports can be highly erratic. Big spikes or other anomalies
can occur in the data as reporting backlogs are cleared, changes in case
definitions are made, etc. Groups like JHU CSSE then work tirelessly to correct
such anomalies after first publication (e.g., they will attempt to
back-distribute a spike when a reporting backlog is cleared, by working with a
local authority to figure out how this should best be done), which can result
in very nontrivial revisions.  See
Figure~\ref{fig:vintage} for an example.

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then.  Furthermore, the
latency of the signal was always at least one week through 2020.  However, this
signal is never revised after initial publication (confirmed via personal
communication with the Google team that produces this signal) and furthermore
the latency of the signal is not an unavoidable property of the data type, so we
simply use finalized signal values, with zero latency, in our analysis.

@@

To fix notation, let $Y_{\ell,t}$ denote the 7-day trailing average of COVID-19
case incidence rates in location (HRR) $\ell$ and at time (day) $t$.  To be
clear, this is the number of new daily reported cases per 100,000
people, averaged over the 7-day period $t-6, \ldots, t$. The first task we
consider---\textit{forecasting}---is to predict $Y_{\ell,t+a}$ for each
``ahead'' value $a=7,\ldots,21$.  The second task---\textit{hotspot
  prediction}---is to predict a binary variable defined in terms of the relative
change of $Y_{\ell,t+a}$ (relative to its value one week prior,
$Y_{\ell,t+a-7}$), again for each $a=7,\ldots,21$.

Why do we define the response variables via 7-day averaging? The short answer 
is robustness: averaging stabilizes the case time series, and accounts for 
uninteresting artifacts like day-of-the-week effects in the series.  Note
that we can equivalently view this (equivalent up to a constant
factor) as predicting the HRR-level COVID-19 case incidence rate \textit{summed}
over some 7-day period in the future, and predicting a binary variable derived
from this.

@@

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data and lagged values
of signals for features, as we will detail shortly) June 9, 2020 for forecasting,
and June 16, 2020 for hotspot prediction. (The bottleneck here is the
CTIS-CLI-in-community signal, which does not exist before early April 2020, when
the survey was first launched). The end date was chosen again with a
consideration to align both tasks as best as possible, and because few hotspots
exist post December 31, 2020, due to the general and gradual decline of the
pandemic in 2021.

@@

Recall $Y_{\ell,t}$ denotes the 7-day trailing average of COVID-19 case
incidence rates in location $\ell$ and at time $t$.  Separately for each
$a=7,\ldots,21$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
simple probabilistic forecasting model of the form:     
\begin{equation}
\label{eq:forecast_ar}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ s \leq t)  
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
\end{equation}
This model uses current case rates, and the case rates 7 and 14 days ago, in
order to predict (the quantiles of) case rates in the future.  We consider a
total of 7 quantile levels (chosen in accordance with the county-level quantile
levels suggested by the COVID-19 Forecast Hub),      
\begin{equation}
\label{eq:quantile_levels}
\tau \in \{0.025,\ 0.1,\ 0.25,\ 0.5,\ 0.75,\ 0.9,\ 0.975 \}.
\end{equation}
We fit \eqref{eq:forecast_ar} using \textit{quantile regression}
\cite{Koenker:1978, Koenker:2005, Koenker:2006} separately for each $\tau$, 
using data from all 306 HRRs, and within each HRR, using the most recent 21 days
of training data.  This gives us 6,426 training samples for each quantile 
regression problem.

In addition to this pure autoregressive model, we also consider five
probabilistic forecasting models of the form:  
\begin{multline}
\label{eq:forecast_ar_x}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t)  
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
\end{multline}
where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
CHNG-COVID, CTIS-CLI-in-community, DV-CLI, or Google-AA---at location $\ell$ and
time $t$. Note that we apply the same lags (current value, along with the values
7 and 14 days ago) for the auxiliary indicators as we do for the case
rates. Training then proceeds just as before: we use the same 7 quantile levels
in \eqref{eq:quantile_levels}, and fit quantile regression separately for each 
level $\tau$, using data from all 306 HRRs and a trailing window of 21 days of
training data.

At prediction time, in order to avoid crossing violations (that is,
for two levels $\tau' > \tau$,
the predicted quantile at level $\tau$ exceeds the predicted quantile at level
$\tau'$), we apply a simple post-hoc sorting.  See Figure~\ref{fig:trajectory}
for an example forecast.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
\caption{This figure displays a forecast from an autoregressive model for the
  HRR containing
  New York City made  
  on October 15 (vertical line).  The fan displays 50\%, 80\% and 95\% intervals
  while the orange curve shows the median forecast. The black curve represents
  ``finalized'' data, as reported in May 2021.} 
\label{fig:trajectory}
\end{figure}

@@

We treat hotspot prediction as a binary classification problem and use a setup 
altogether quite similar to the forecasting setup described previously.
Separately for each $a=7,\ldots,21$, to predict $Z_{\ell,t+a}$, we consider a
simple logistic model:
\begin{equation}
\label{eq:hotspot_ar}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ s \leq t) \big) 
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
\end{equation}
where $\logit(p) = \log(p/(1-p))$, the log-odds of $p$.

In addition to this pure autoregressive model, we also consider five
logistic models of the form:  
\begin{multline}
\label{eq:hotspot_ar_x}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t) \big)
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit the above models,
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x}, using logistic regression,   
pooling all 306 HRRs and using a 21-day trailing window for the training data.

@@

For forecasting, we evaluate the probabilistic forecasts produced by the
quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using 
\textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
\cite{Gneiting:2007}.  WIS is a proper score, meaning that its expectation
is minimized by the population quantiles of the target variable.  The use of WIS
in COVID-19 forecast scoring is discussed in \cite{Bracher:2021}; WIS is also
the main evaluation metric used in the COVID-19 Forecast Hub.

WIS is typically defined for quantile-based forecasts where the quantile levels
are symmetric around 0.5.  This is the case for our choice in
\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the case of
symmetric quantile levels, this is equivalent to a collection of central
prediction intervals $(\ell_\alpha, u_\alpha)$, parametrized by an exclusion
probability $\alpha$. The WIS of the forecaster $F$, evaluated at the target
variable $Y$, is defined by:
\begin{equation}
\label{eq:wis_intervals}
\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
\end{equation}
where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
smallest distance between $a$ and an element of $S$).  Note that, corresponding 
to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
\{0.05,\ 0.2,\ 0.5,\ 1\}$, resulting in 4 terms in the above sum.  By 
straightforward algebra, it is not hard to see WIS has an alternative 
representation in terms of the predicted quantiles themselves:
\begin{equation}
\label{eq:wis_quantiles}
\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
\end{equation}
where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
presentation, because the score nicely decouples into a ``sharpness'' component
(first term in each summand) and an ``under/overprediction'' component (second term
in each summand).  The second form given in \eqref{eq:wis_quantiles} is
especially noteworthy in our current study because it reveals WIS to be the same as the
quantile regression loss that we use to train our forecasting models (i.e., our
models are fit to optimize WIS averaged over the training set).

For hotspot prediction, we evaluate the probabilistic classifiers produced by
the logistic models in \eqref{eq:hotspot_ar} and \eqref{eq:hotspot_ar_x} using the 
area under the curve (AUC) of their true positive versus false positive rate
curve (which is traced out by varying the discrimination threshold).

The primary aggregation scheme that we will use in model evaluation and
comparisons will be to average WIS per forecaster at ahead value $a$ over all
forecast dates $t$ and locations $\ell$; and similarly, to compute AUC per
classifier at ahead value $a$ over all forecast dates $t$ and locations
$\ell$.
